%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{Acknowledgments} % Bookmark name visible in a PDF viewer

\chapter*{Acknowledgments}

In this complex project, many people provided advice, assistance and support
during the past six years. I remember gratefully:

For guidance in accomplishing this work, my advisor Alexandre Perera Lluna.
I appreciate a lot your support, guidance and openness to share your knowledge with me.
It has meant so much to me over the years, both professionally and personally. 

For opportunity to do a part of my work in their laboratories,
Santiago Marco and laboratory of the Intelligent Signal Processing (ISP) group 
at University of Barcelona,
Paul Verschure and laboratory of the Synthetic Perceptive, Emotive and Cognitive Systems (SPECS) group
at Pompeu Fabra University, and 
Romeo Beccherelli and laboratory of Consiglio Nazionale delle Ricerche (CNR)
at Institute for Microelectronics and Microsystems.
For giving access to the unique UNIMAN data set, 
Professor at the University of Manchester Krishna Persaud.

For working in close co-operation in the \Neurochem project,
Jordi Fonollosa, Luis Fern\'{a}ndez, Miguel Lech\'{o}n,
Sergi Berm\'{u}dez Badia, 
Lucas L\'{o}pez, Vasiliki Vouloutsi, Alex Chimeno Escudero
and Emiliano Zampetti.

For the department-level support, 
director of the Biomedical Engineering Research Centre (CREB) Pere Caminal, and 
director of the PhD Program in Bioengineering Raimon Jan\'{e}.
For providing painless maintenance of the IT infrastructure, 
Guillem Riera Galm\'{e}s.
For coping with the paper work, Virginia Cabello and Marta Clar\'{o}s.
For studying and working together in the ESAII department, 
Helena Brunel, Juan Maynou, Raimon Massanet, Joan Josep Gallardo, Francesc Fern\'{a}ndez,
Jorge Jaramillo, Joan Francesc Alonso, M\'{o}nica Rojas, Umberto Melia and Leidy Serna.

For investing time and effort in revision of this manuscript, Oleg Osychenko. 
For translating the abstract of this manuscript from English to Spanish, Helena Brunel.
For sharing with me their thesis manuscripts as a reference
of the format of a collection of published articles,
Yorgos Athanasiadis, Joan Francesc Alonso and Narges Dialami.

For funding my PhD study,
the European Community's Seventh Framework Programme (FP7/2007-2013)
under grant agreement no. 216916: 
Biologically inspired computation for chemical sensing (Neurochem).
For giving me the opportunity to work at 
the Institute of Biomedical Research of Hospital de la Santa Creu i Sant Pau
for the last two years of my study,
head of the Unit of Genomics of Complex Diseases Jos\'{e} Manuel Soria,
and head of the Unit of Thrombosis and Haemostasis Juan Carlos Souto.

For open science and open-source software projects related to my research, 
the Debian developers, the Comedi project developers,
the R core development team, many creators of R packages, 
the maintainers of The University of California at Irvine (UCI) 
Machine Learning Repository, and
the classicthesis package for \LaTeX.

For being on the other side of my PhD study, friends I met in Barcelona
Maxim, Tanya, Eduardo, Martin, Israel, Mariella, Giovanna, 
Oleg, Lourdes, Gregory, Eva, Katya, Marina, Pavel and Narges.

For absolute confidence in me and care, my mother Lubov, my sister Lubava and my grandmother Raisa.
For being together with me, Nina.

